Advanced Methods for short-term Path Planning in Autonomous Driving based on Adaptive MPC

### 2019-01-09 Thesis_Latex/Paper/Paper.tex
update MATLAB files and PAPER

\title{Advanced Methods for Autonomous Driving based on Adaptive Model Predictive Control}
\author{\IEEEauthorblockN{Alberto Franco}\\
\IEEEauthorblockA{Department of Information Engineering\\
Università degli Studi di Padova\\ \href{mailto:alberto.franco.3@studenti.unipd.it}{alberto.franco.3@studenti.unipd.it}}}

\begin{document}
\maketitle
\begin{abstract}
	In this paper we present two different strategies for a self-driving car. The aim of this project is to study and implement a path tracking framework in order to prevent ATLASCAR2 collision with moving obstacle vehicles. The proposed algorithms, based on the Model Predictive Control paradigm, solve optimal problems formulated in terms of cost minimization under constraints.
\end{abstract}

@@

\section{Context of the Problem and \\  Proposed Solution}
The ATLACAR2 has a sensor, such as lidar, that measures the distance to an obstacle in front of the vehicle and in the same lane. The obstacle can be static, such as a large pothole, or moving, such as a slow-moving vehicle. The most common maneuver from the driver is to temporarily move to another lane, drive past the obstacle, and move back to the original lane afterward.
In this case, we want to design an obstacle avoidance system that moves the ATLASCAR2 around a moving obstacle in the lane using throttle and steering angle. This system uses an adaptive Model Predictive Controller that updates both the predictive model and the mixed input/output constraints at each control interval. Moreover this vehicle is also 
equipped with a lane-keeping assist system: it has a sensor, such as camera, that measures the lateral deviation and relative yaw angle between the centerline of a lane and the ATLASCAR2. The sensor also measures the current lane curvature and curvature derivative. Depending on the curve length that the sensor can view, the curvature in front of the vehicle can be calculated from the current curvature and curvature derivative. This system keeps the autonomous car travelling along the centerline of the lanes on the road by adjusting the front steering angle. The goal for lane keeping control is to drive both lateral deviation and relative yaw angle close to zero.
\subsection{Theoretical background - adaptive Model Predictive Control}
Model Predictive Control is an advanced method that predicts future behavior using a linear-time-invariant (LTI) dynamic model. These predictions are never exact and a good strategy is
to make MPC insensitive to prediction errors. If the plant is strongly nonlinear or its characteristics vary dramatically with time, MPC performance might become unacceptable because LTI prediction accuracy degrade. A method that can address this degradation by adapting the prediction model for changing operating conditions is called Adaptive MPC: this control strategy uses a fixed model structure, but allows the models parameters to evolve with time. Ideally, whenever the controller requires a prediction, it uses a model appropriate for the current conditions. At each control interval, the adaptive MPC controller updates the plant model and nominal conditions. Once updated, the model and conditions remain constant over the prediction horizon. The plant model used as the basis for the adaptive MPC must be an LTI discrete-time, state-space model with a structure as follows:
\begin{equation*}
\label{eqn:Adaptive_MPC_plant_discrete}
\begin{aligned}
x(k+1)&=Ax(k)+ B_u u(k)+B_v v(k)+B_d d(k)\\
y(k)&=Cx(k) + D_v v(k)+ D_d d(k)
\end{aligned}
\end{equation*}
where the matrices $A$, $B_u$, $B_v$, $B_d$, $C$, $D_v$ and $D_d$ can vary with time. The other parameters in the expression are:
\begin{itemize}
	\item $k$ is the time index / current control interval;
	\item $x$ are the plant model states;
	\item $u$ are the manipulated inputs that can be adjusted by the MPC controller;
	\item $v$ are the measured disturbance inputs;
	\item $d$ are the unmeasured disturbance inputs;
	\item $y$ are the plant outputs, including both measured and unmeasured. Is it necessary at least one measured output.
\end{itemize}

@@

\section{Obstacle Avoidance}
\subsection{Problem Formulation}
The collision avoidance problem is very dependent on the vehicle modeling since it is a requirement for  adaptive MPC law design. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=1\columnwidth]{./figure/obstacleAvoidance.pdf}
	\caption{Problem description of collision avoidance on road.}
	\label{fig:obstacleAvoidance}
\end{figure}

@@

The simple non-linear model to describe the dynamics of the ATLASCAR2 is the following
\begin{equation*}
	\label{eqn:dynamics_model_obstacle_avoidance}
	\left \{ \begin{array}{llll}
	\dot{x} = -v\sin(\theta)\cdot\theta+\cos(\theta)\cdot v\\
	\dot{y} = v\cos(\theta)\cdot\theta+\sin(\theta)\cdot v\\
	\dot{\theta} =(\tan(\delta)/C_L)\cdot v+ \bigg(v\bigg(\tan(\delta)^2+1\bigg)/C_L\bigg)\cdot \delta\\
	\dot{v} =0.5\cdot T
	\end{array} 
	\right .
\end{equation*}
where $C_L$ is the car length. The state equations can be re-written in a more compact form as follows:
\begin{equation}
	\label{eqn:dynamics_ss_obstacle_avoidance_con}
	\begin{array}{ll}
	\dot{\textbf{x}} =A_c \textbf{x}+ B_c \textbf{u}\\
	\textbf{y} =C_c \textbf{x} + D_c \textbf{u}
	\end{array}
\end{equation}
where
\[ 
A_c=\begin{bmatrix}
0&0&-v\sin(\theta)&\cos(\theta)\\
0&0&v\cos(\theta)&\sin(\theta)\\
0&0&0&\tan(\delta)/C_L\\
0&0&0&0
\end{bmatrix},
\]
\[
B_c=\begin{bmatrix}
0&0\\
0&0\\
0&v\bigg(\tan(\delta)^2+1\bigg)/C_L\\
0.5&0
\end{bmatrix},
\]
\[
C_c=\begin{bmatrix}
1&0&0&0\\
0&1&0&0\\
0&0&1&0\\
0&0&0&1
\end{bmatrix} = \textbf{I}_4, 
\qquad
D_c=\begin{bmatrix}
0&0\\
0&0\\
0&0\\
0&0
\end{bmatrix} = \textbf{0}_{4\times2}. 
\]
The system to be controlled is usually modeled by a discrete state-space model in the MPC literature. Therefore, (\ref{eqn:dynamics_ss_obstacle_avoidance_con})
is transformed into a discrete state-space model to be used by the model predictive controller:
\begin{equation}
\label{eqn:dynamics_ss_obstacle_avoidance_dis}
	\begin{array}{rr}
	{\textbf{x}}(k+1) =A_d \textbf{x}(k)+ B_d \textbf{u}(k)\\
	\textbf{y}(k) =C_d \textbf{x}(k) + D_d \textbf{u}(k)
	\end{array}
\end{equation}
where $A_d$ and $B d$ are the state and control matrices for the discrete state-space equation, respectively, which can be calculated with the Euler method as
\[
A_d = e^{A_cT_s},\qquad B_d = \int_{kT_s}^{(k+1)T_s} e^{A_c[(k+1)T_s-\eta]}B_c d\eta
\]
where $T_s$ is the sampling interval for the discrete state-space model. The matrices $C_d$ and $D_d$ are equivalent to those in the continuous case. For simplicity, we assume that all the states are measurable and the ATLASCAR2 drives east with a constant speed at the nominal operating point. In the scenario we are going to consider, the road is straight and our vehicle stays in the middle of the center lane when not passing. Without losing generality, the ATLASCAR2 passes an obstacle both to the right and to the left lane depending on where it is placed on the road. We create also a safe zone around the obstacles so that the vehicle does not get too close the osbtacle when passing it.

@@

We designed a model predictive controller that can make the ATLASCAR2 mantain a desired velocity and stay in the middle of center lane. We used an Adaptive MPC controller because it handles the nonlinear vehicle dynamics more effectively than a traditional MPC controller in fact the latter uses a constant plant model, however, the former allows us to provide a new plant model at each control interval. Because the new model describes the plant dynamics more accurately at the new operating condition, an adaptive MPC controller performs better than a traditional MPC controller. In practice at each control interval, the adaptive MPC controller updates the plant model and the nominal conditions. Once updated, the model and the conditions remain constant over the prediction horizon. At the beginning, we specified the constraints for the manipulated variables: to prevent the ATLASCAR2 from accelerating or decelerating too quickly, we added an hard constraint on the throttle rate of change and another one on the steering angle rate of change. We used an approach that takes advantage of the ability of MPC to handle constraint explicitly. When an obstacle is detected, it defines an area on the road (in terms of constraints) that the ATLASCAR2 must not enter during the prediction horizon. At the next control interval the area is redefined based on the new positions of the vehicle and the obstacle until passing is completed.
To define the area to avoid, we used the following mixed Input/Output constraints:
\begin{equation*}
	\label{eqn:mixed_IO_constraints}
	E\textbf{u}+F\textbf{y}\leq G
\end{equation*}
where $\textbf{u}$ and $\textbf{y}$ are respectively the manipulated variable vector and the output variable vector, while $E,F,G$ are the matrices that can be updated when the controller is running:
\begin{itemize}
	\item the first constraint is an upper bound on the $y$-coordinate (right boundary of the road);
	\item the second constraint is a lower bound on the $y$-coordinate (left boundary of the road);
	\item the third constraint is for the obstacle avoidance; even though no obstacle is detected at the nominal condition, we must add this fake constraint here because we cannot change the dimensions of the constraint matrices at run time (it has the same form as the second one).
\end{itemize}
The matrices for the above inequality are the following:
\[
E= 
\begin{bmatrix}
0&0\\
0&0\\
0&0\\
0&0
\end{bmatrix},
\quad
F=\begin{bmatrix}
0&1&0&0\\
0&-1&0&0\\
cS&-1&0&0
\end{bmatrix},\quad
G=
\begin{bmatrix}
W/2\\W/2\\-cI
\end{bmatrix}
\]
where $W$ is the width of the road and $cI$ and $cS$ are the required parameters such that the ATLASCAR2 must be above the line formed from the vehicle to safe zone corner for left/right passing.

@@

\section{Lane Following}
\subsection{Problem Formulation}
A lane-following system is a control system that keeps the vehicle traveling along the centerline of a highway lane, while maintaining a user-set velocity. The lane-following scenario is depicted in the following figure:
\begin{figure}[!h]
	\centering
	\includegraphics[width=1\columnwidth]{./figure/laneFollowing.pdf}
	\caption{Problem description of a lane following system.}
	\label{fig:laneFollowing}
\end{figure}

In a classic lane keeping assist, it is assumed that the longitudinal velocity is constant. This restriction is relaxed in this model because the longitudinal acceleration varies in this MIMO control system. This lane-following system manipulates both the longitudinal acceleration and the front steering angle of the vehicle to keep the lateral deviation and the relative yaw angle small and the longitudinal velocity close to a driver set velocity. If these two goals cannot be met at the same moment, the system tries to balance them. The model that we are considering, contains a lot of parameters. The first fundamental block describes the vehicle dynamics: we have applied the bicycle model of lateral vehicle dynamics and approximate the longitudinal dynamics using a time constant:
\subsubsection{Longitudinal dynamics}
\begin{equation}
	\label{eqn:longi_dynamics_simple_model}
	G(s)=\frac{1}{\tau s^2+s} \quad \text{where $\tau$ is a time constant}
\end{equation}

\subsubsection{Lateral dynamics}
Local function: we have a continuous vehicle lateral model from parameters
\begin{equation}
\label{eqn:lateral_dynamics_simple_model}
\begin{array}{ll}
\dot{\textbf{x}}_{\text{lat}} =A_g \textbf{x}_{\text{lat}}+ B_g \textbf{u}_{\text{lat}}\\
\textbf{y}_{\text{lat}} =C_g \textbf{x}_{\text{lat}} + D_g \textbf{u}_{\text{lat}}
\end{array}
\end{equation}
where the input is the steering angle in radians, and the outputs are the lateral velocity in meters per second and yaw angle rate in radians per second:
\begin{equation*}
\textbf{x}_{\text{lat}} = \begin{bmatrix}
V_y&\dot{\psi}
\end{bmatrix}^T
\qquad
\textbf{u}_{\text{lat}} = \delta
\end{equation*}
and
\[ 
A_g=\begin{bmatrix}
-\frac{2C_F+2C_R}{mV_x}&-\frac{2C_Fl_F-2C_Rl_R}{mV_x} - V_x\\
-\frac{2C_Fl_F-2C_Rl_R}{I_ZV_x}&-\frac{2C_Fl_F^2+2C_Rl_R^2}{I_ZV_x}
\end{bmatrix},
\]
\[
B_g=\begin{bmatrix}
2C_F/m\\
2C_Fl_F/I_Z
\end{bmatrix},
\quad
C_g=\textbf{I}_2, 
\quad
D_g=\textbf{0}_{2\times1}.
\]
The parameters in the previous matrices are:
\begin{itemize}
	\item $V_x$ is the longitudinal velocity of the car;	
	\item $m$ is the total mass parameter; 
	\item $I_Z$ is the yaw moment of inertia parameter;
	\item $l_F$ and $l_R$ are the longitudinal distances from center of gravity to front and rear tires parameters;
	\item $C_F$ and $C_R$ are the cornering stiffnesses of front and rear tires parameters.
\end{itemize}
The goal for the driver steering model is to keep the vehicle in its lane and follow the curved road by controlling the front steering angle . This goal is achieved by driving the yaw angle error $e_2 = \psi -\psi_{\text{des}}$ and lateral displacement error $e_1$ to zero ($\dot{e}_1 = V_xe_2+V_y$). We can incorporate these two paramenters in the augmented model:
\begin{equation}
\label{eqn:lateral_dynamics_augmented_model}
\begin{array}{ll}
\dot{\textbf{x}}_{\text{aug}} =A_a \textbf{x}_{\text{aug}}+ B_a \textbf{u}_{\text{aug}}\\
\textbf{y}_{\text{aug}} =C_a \textbf{x}_{\text{aug}} + D_a \textbf{u}_{\text{aug}}
\end{array}
\end{equation}
where
\begin{equation*}
\textbf{x}_{\text{aug}} = \begin{bmatrix}
V_y&\dot{\psi}&e_1&e_2
\end{bmatrix}^T,
\qquad
\textbf{u}_{\text{aug}} = 
\begin{bmatrix}
\delta&\dot{\psi}_{\text{des}}
\end{bmatrix}^T
\end{equation*}
and
\[ 
A_a=\begin{bmatrix}
A_g&\textbf{0}_{2\times2}\\
\textbf{I}_2&\begin{matrix}
0&V_x\\
0&0
\end{matrix}
\end{bmatrix},
\quad
B_a=\begin{bmatrix}
B_g&\textbf{0}_{2\times1}\\
0&0\\
0&-1
\end{bmatrix},
\]
\[
C_a=\begin{bmatrix}
\textbf{0}_{2\times2}&\textbf{I}_2
\end{bmatrix}, 
\quad
D_a=
\textbf{0}_{2\times2}. 
\]
Combining (\ref{eqn:longi_dynamics_simple_model}) with (\ref{eqn:lateral_dynamics_augmented_model}) yields the state-space model that characterizes the model predictive controller:
\begin{equation}
\label{eqn:full_dynamics_model}
\begin{array}{ll}
\dot{\textbf{x}}_{\text{tot}} =A_f \textbf{x}_{\text{tot}}+ B_f \textbf{u}_{\text{tot}}\\
\textbf{y}_{\text{tot}} =C_f \textbf{x}_{\text{tot}} + D_f \textbf{u}_{\text{tot}}
\end{array}
\end{equation}
where
\begin{equation*}
\textbf{x}_{\text{tot}} = \begin{bmatrix}
V_x\\\dot{V}_x\\V_y\\\dot{\psi}\\e_1\\e_2
\end{bmatrix},
\qquad
\textbf{u}_{\text{tot}} = 
\begin{bmatrix}
a\\\delta\\\dot{\psi}_{\text{des}}
\end{bmatrix}
\end{equation*}
and

\[ 
A_f=\begin{bmatrix}
\begin{matrix}-2&0\\1&0\end{matrix}&\textbf{0}_{2\times4}\\
\textbf{0}_{4\times2}&A_a
\end{bmatrix},
\quad
B_f=\begin{bmatrix}\begin{matrix}2\\0\end{matrix}&\textbf{0}_{2\times2}\\
\textbf{0}_{4\times1}&B_a
\end{bmatrix},
\]
\[
C_f=\begin{bmatrix}
\begin{matrix}0&1\end{matrix}&\textbf{0}_{1\times4}\\
\textbf{0}_{2\times2}&C_a
\end{bmatrix}, 
\quad
D_f=\textbf{0}_{3\times3}. 
\]

@@

@@

>>>>

\title{Advanced Methods for short-term Path Planning in Autonomous Driving based on Adaptive MPC}
\author{\IEEEauthorblockN{Alberto Franco}\\
\IEEEauthorblockA{Department of Information Engineering\\
Università degli Studi di Padova\\ \href{mailto:alberto.franco.3@studenti.unipd.it}{alberto.franco.3@studenti.unipd.it}}}

\begin{document}
\maketitle
\begin{abstract}
	In this paper we present two different strategies for a self-driving car short-term path planning. The aim of this project is to study and implement a path tracking framework in order to prevent ATLASCAR2 collision with moving obstacle vehicles. The proposed algorithms, based on the Model Predictive Control paradigm, solve optimal problems formulated in terms of cost minimization under constraints.
\end{abstract}

@@

\section{Context of the Problem and \\  Proposed Solution}
The ATLACAR2 has sensors, such as lidar, that measure the distance to obstacles in front and around the vehicle. The obstacles can be static, such as a large pothole, or moving, such as a slow-moving vehicle. The most common maneuver from the driver is to temporarily move to another lane, drive past the obstacle, and move back to the original lane afterward.
In this case, we want to design an obstacle avoidance system that moves the ATLASCAR2 around a moving obstacle in the lane using throttle and steering angle. This system uses an adaptive Model Predictive Controller that updates both the predictive model and the mixed input/output constraints at each control interval. Moreover, this vehicle is also equipped with a lane-keeping assist system: it has a sensor, such as camera, that measures the lateral deviation and relative yaw angle between the centerline of a lane and the ATLASCAR2. The sensor also measures the current lane curvature and curvature derivative. Depending on the curve length that the sensor can view, the curvature in front of the vehicle can be calculated from the current curvature and curvature derivative. This system keeps the autonomous car travelling along the centerline of the lanes on the road by adjusting the front steering angle. The goal for lane keeping control is to drive both lateral deviation and relative yaw angle close to zero.
\subsection{Theoretical background - adaptive Model Predictive Control}
Model Predictive Control is an advanced method that predicts future behavior using a linear-time-invariant (LTI) dynamic model. These predictions are never exact and a good strategy is to make MPC insensitive to prediction errors. If the plant is strongly nonlinear or its characteristics vary dramatically with time, MPC performance might become unacceptable because LTI prediction accuracy degrade. A method that can address this degradation by adapting the prediction model for changing operating conditions is called Adaptive MPC: this control strategy uses a fixed model structure, but allows the model parameters to evolve with time. Ideally, whenever the controller requires a prediction, it uses a model appropriate for the current conditions. At each control interval, the adaptive MPC controller updates the plant model and nominal conditions. Once updated, the model and conditions remain constant over the prediction horizon. The plant model used as the basis for the adaptive MPC must be an LTI discrete-time, state-space model with a structure as follows:
\begin{equation*}
\label{eqn:Adaptive_MPC_plant_discrete}
\begin{aligned}
x(k+1)&=Ax(k)+ B_u u(k)+B_v v(k)+B_d d(k)\\
y(k)&=Cx(k) + D_v v(k)+ D_d d(k)
\end{aligned}
\end{equation*}
where the matrices $A$, $B_u$, $B_v$, $B_d$, $C$, $D_v$ and $D_d$ can vary with time. The other parameters in the expression are:
\begin{itemize}
	\item $k$ is the time index / current control interval;
	\item $x$ are the plant model states;
	\item $u$ are the manipulated inputs that can be adjusted by the MPC controller;
	\item $v$ are the measured disturbance inputs;
	\item $d$ are the unmeasured disturbance inputs;
	\item $y$ are the plant outputs, including both measured and unmeasured. Is it necessary at least one measured output.
\end{itemize}

@@

\section{Obstacle Avoidance}
\subsection{Problem Formulation}
The collision avoidance problem is very dependent on the vehicle modeling since it is a requirement for  adaptive MPC law design. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=1\columnwidth]{./figure/obstacleAvoidance/obstacleAvoidance.pdf}
	\caption{Problem description of collision avoidance on road.}
	\label{fig:obstacleAvoidance}
\end{figure}

@@

The simple non-linear model to describe the dynamics of the ATLASCAR2 is derived combining the equations in [2] and [3]:
\begin{equation*}
	\label{eqn:dynamics_model_obstacle_avoidance}
	\left \{ \begin{array}{llll}
	\dot{x} = -v\sin(\theta)\cdot\theta+\cos(\theta)\cdot v\\
	\dot{y} = v\cos(\theta)\cdot\theta+\sin(\theta)\cdot v\\
	\dot{\theta} =(\tan(\delta)/C_L)\cdot v+ \bigg(v\bigg(\tan(\delta)^2+1\bigg)/C_L\bigg)\cdot \delta\\
	\dot{v} =0.5\cdot T
	\end{array} 
	\right .
\end{equation*}
where $C_L$ is the car length. The state equations can be re-written in a more compact form as follows:
\begin{equation}
	\label{eqn:dynamics_ss_obstacle_avoidance_con}
	\begin{array}{ll}
	\dot{\textbf{x}} =A_c \textbf{x}+ B_c \textbf{u}\\
	\textbf{y} =C_c \textbf{x} + D_c \textbf{u}
	\end{array}
\end{equation}
where
\[ 
A_c=\begin{bmatrix}
0&0&-v\sin(\theta)&\cos(\theta)\\
0&0&v\cos(\theta)&\sin(\theta)\\
0&0&0&\tan(\delta)/C_L\\
0&0&0&0
\end{bmatrix},
\]
\[
B_c=\begin{bmatrix}
0&0\\
0&0\\
0&v\bigg(\tan(\delta)^2+1\bigg)/C_L\\
0.5&0
\end{bmatrix},
\]
\[
C_c=\begin{bmatrix}
1&0&0&0\\
0&1&0&0\\
0&0&1&0\\
0&0&0&1
\end{bmatrix} = \textbf{I}_4, 
\qquad
D_c=\begin{bmatrix}
0&0\\
0&0\\
0&0\\
0&0
\end{bmatrix} = \textbf{0}_{4\times2}. 
\]
The system to be controlled is usually modeled by a discrete state-space model in the MPC literature. Therefore, (\ref{eqn:dynamics_ss_obstacle_avoidance_con})
is transformed into a discrete state-space model to be used by the model predictive controller:
\begin{equation}
\label{eqn:dynamics_ss_obstacle_avoidance_dis}
	\begin{array}{rr}
	{\textbf{x}}(k+1) =A_d \textbf{x}(k)+ B_d \textbf{u}(k)\\
	\textbf{y}(k) =C_d \textbf{x}(k) + D_d \textbf{u}(k)
	\end{array}
\end{equation}
where $A_d$ and $B d$ are the state and control matrices for the discrete state-space equation, respectively, which can be calculated with the Euler method as
\[
A_d = e^{A_cT_s},\qquad B_d = \int_{kT_s}^{(k+1)T_s} e^{A_c[(k+1)T_s-\eta]}B_c d\eta
\]
where $T_s$ is the sampling interval for the discrete state-space model. The matrices $C_d$ and $D_d$ are equivalent to those in the continuous case. For simplicity, we assume that all the states are measurable and the ATLASCAR2 drives east with a constant speed at the nominal operating point. In the scenario that we are going to consider, the road is straight and our vehicle stays in the middle of the center lane when not passing. Without losing generality, the ATLASCAR2 passes an obstacle both to the right and to the left lane depending on where it is placed on the road. We create also a safe zone around the obstacles so that the vehicle does not get too close to the osbtacle when passing it.

@@

We designed a model predictive controller that can make the ATLASCAR2 mantain a desired velocity and stay in the middle of center lane. We used an Adaptive MPC controller because it handles the nonlinear vehicle dynamics more effectively than a traditional MPC controller; in fact, the latter uses a constant plant model but the former allows us to provide a new plant model at each control interval. Because the new model describes the plant dynamics more accurately at the new operating condition, an adaptive MPC controller performs better than a traditional MPC controller. In practice at each control interval, the adaptive MPC controller updates the plant model and the nominal conditions. Once updated, the model and the conditions remain constant over the prediction horizon. At the beginning, we specified the constraints for the manipulated variables: to prevent the ATLASCAR2 from accelerating or decelerating too quickly, we added an hard constraint on the throttle rate of change and another one on the steering angle rate of change. We used an approach that takes advantage of the ability of MPC to handle constraint explicitly. When an obstacle is detected, it defines an area on the road (in terms of constraints) that the ATLASCAR2 must not enter during the prediction horizon. At the next control interval the area is redefined based on the new positions of the vehicle and the obstacle until passing is completed.
To define the area to avoid, we used the following mixed Input/Output constraints:
\begin{equation*}
	\label{eqn:mixed_IO_constraints}
	E\textbf{u}+F\textbf{y}\leq G
\end{equation*}
where $\textbf{u}$ and $\textbf{y}$ are respectively the manipulated variable vector and the output variable vector, while $E,F,G$ are the matrices that can be updated when the controller is running:
\begin{itemize}
	\item the first constraint is an upper bound on the $y$-coordinate (right boundary of the road);
	\item the second constraint is a lower bound on the $y$-coordinate (left boundary of the road);
	\item the third constraint is for the obstacle avoidance; even though no obstacle is detected at the nominal condition, we must add this fake constraint here because we cannot change the dimensions of the constraint matrices at run time (it has the same form as the second one).
\end{itemize}
The matrices for the above inequality are the following:
\[
E= 
\begin{bmatrix}
0&0\\
0&0\\
0&0\\
0&0
\end{bmatrix},
\quad
F=\begin{bmatrix}
0&1&0&0\\
0&-1&0&0\\
cS&-1&0&0
\end{bmatrix},\quad
G=
\begin{bmatrix}
W/2\\W/2\\-cI
\end{bmatrix}
\]
where $W$ is the width of the road and $cI$ and $cS$ are the required parameters such that the ATLASCAR2 must be above the line formed from the vehicle to safe zone corner for left/right passing.

@@

\section{Lane Following}
\subsection{Problem Formulation}
A lane-following system is a control system that keeps the vehicle traveling along the centerline of a highway lane, while maintaining a user-set velocity. The lane-following scenario is depicted in the following figure:
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.75\columnwidth]{./figure/laneFollowing/laneFollowing.pdf}
	\caption{Problem description of a lane following system.}
	\label{fig:laneFollowing}
\end{figure}

In a classic lane keeping assist, it is assumed that the longitudinal velocity is constant. This restriction is relaxed in this model because the longitudinal acceleration varies in this MIMO control system. This lane-following system manipulates both the longitudinal acceleration and the front steering angle of the vehicle to keep the lateral deviation and the relative yaw angle small and the longitudinal velocity close to a driver set velocity. If these two goals cannot be met at the same moment, the system tries to balance them. The model that we are considering, contains a lot of parameters. The first fundamental block describes the vehicle dynamics: we have applied the bicycle model of lateral vehicle dynamics and approximate the longitudinal dynamics using a time constant:
\subsubsection{Longitudinal dynamics}
We can use the following state space to describe the longitudinal model:
\begin{equation}
\label{eqn:longi_dynamics_simple_model_ss}
\begin{array}{ll}
\dot{\textbf{x}}_{\text{lon}} =A_m \textbf{x}_{\text{lon}}+ B_m\textbf{u}_{\text{lon}}\\
\textbf{y}_{\text{lon}} =C_m \textbf{x}_{\text{lon}} + D_m \textbf{u}_{\text{lon}}
\end{array}
\end{equation}
where the input is the acceleration and the states are the longitudinal velocity and the actual acceleration which is also the only output of this system.
\begin{equation*}
\textbf{x}_{\text{lon}} = \begin{bmatrix}
\dot{V}_x&V_x
\end{bmatrix}^T,
\qquad
\textbf{u}_{\text{lon}} = a
\end{equation*}
and
\[ 
A_g=\begin{bmatrix}
-\frac{1}{\tau}&0\\1&0
\end{bmatrix},
\quad
B_g=\begin{bmatrix}
\frac{1}{\tau}\\
0
\end{bmatrix},
\]
\[
C_g=\begin{bmatrix}
1&0
\end{bmatrix}, 
\quad
D_g=0.
\]
where $\tau$ is a time constant; in practice we are considering a second order transfer function like in [5].
\subsubsection{Lateral dynamics}
Local function: we have a continuous vehicle lateral model from parameters obtained by simplifying the one in [7]: 
\begin{equation}
\label{eqn:lateral_dynamics_simple_model}
\begin{array}{ll}
\dot{\textbf{x}}_{\text{lat}} =A_g \textbf{x}_{\text{lat}}+ B_g \textbf{u}_{\text{lat}}\\
\textbf{y}_{\text{lat}} =C_g \textbf{x}_{\text{lat}} + D_g \textbf{u}_{\text{lat}}
\end{array}
\end{equation}
where the input is the steering angle in radians, and the outputs are the lateral velocity in meters per second and yaw angle rate in radians per second:
\begin{equation*}
\textbf{x}_{\text{lat}} = \begin{bmatrix}
V_y&\dot{\psi}
\end{bmatrix}^T
\qquad
\textbf{u}_{\text{lat}} = \delta
\end{equation*}
and
\[ 
A_g=\begin{bmatrix}
-\frac{2C_F+2C_R}{mV_x}&-\frac{2C_Fl_F-2C_Rl_R}{mV_x} - V_x\\
-\frac{2C_Fl_F-2C_Rl_R}{I_ZV_x}&-\frac{2C_Fl_F^2+2C_Rl_R^2}{I_ZV_x}
\end{bmatrix},
\]
\[
B_g=\begin{bmatrix}
2C_F/m\\
2C_Fl_F/I_Z
\end{bmatrix},
\quad
C_g=\textbf{I}_2, 
\quad
D_g=\textbf{0}_{2\times1}.
\]
The parameters in the previous matrices are:
\begin{itemize}
	\item $V_x$ is the longitudinal velocity of the car;	
	\item $m$ is the total mass parameter; 
	\item $I_Z$ is the yaw moment of inertia parameter;
	\item $l_F$ and $l_R$ are the longitudinal distances from center of gravity to front and rear tires parameters;
	\item $C_F$ and $C_R$ are the cornering stiffnesses of front and rear tires parameters.
\end{itemize}
The goal for the driver steering model is to keep the vehicle in its lane and follow the curved road by controlling the front steering angle . This goal is achieved by driving the yaw angle error $e_2 = \psi -\psi_{\text{des}}$ and lateral displacement error $e_1$ to zero ($\dot{e}_1 = V_xe_2+V_y$). We can incorporate these two paramenters in the augmented model:
\begin{equation}
\label{eqn:lateral_dynamics_augmented_model}
\begin{array}{ll}
\dot{\textbf{x}}_{\text{aug}} =A_a \textbf{x}_{\text{aug}}+ B_a \textbf{u}_{\text{aug}}\\
\textbf{y}_{\text{aug}} =C_a \textbf{x}_{\text{aug}} + D_a \textbf{u}_{\text{aug}}
\end{array}
\end{equation}
where
\begin{equation*}
\textbf{x}_{\text{aug}} = \begin{bmatrix}
V_y&\dot{\psi}&e_1&e_2
\end{bmatrix}^T,
\qquad
\textbf{u}_{\text{aug}} = 
\begin{bmatrix}
\delta&\dot{\psi}_{\text{des}}
\end{bmatrix}^T
\end{equation*}
and
\[ 
A_a=\begin{bmatrix}
A_g&\textbf{0}_{2\times2}\\
\textbf{I}_2&\begin{matrix}
0&V_x\\
0&0
\end{matrix}
\end{bmatrix},
\quad
B_a=\begin{bmatrix}
B_g&\textbf{0}_{2\times1}\\
0&0\\
0&-1
\end{bmatrix},
\]
\[
C_a=\begin{bmatrix}
\textbf{0}_{2\times2}&\textbf{I}_2
\end{bmatrix}, 
\quad
D_a=
\textbf{0}_{2\times2}. 
\]
Combining (\ref{eqn:longi_dynamics_simple_model_ss}) with (\ref{eqn:lateral_dynamics_augmented_model}) yields the state-space model that characterizes the model predictive controller:
\begin{equation}
\label{eqn:full_dynamics_model}
\begin{array}{ll}
\dot{\textbf{x}}_{\text{tot}} =A_f \textbf{x}_{\text{tot}}+ B_f \textbf{u}_{\text{tot}}\\
\textbf{y}_{\text{tot}} =C_f \textbf{x}_{\text{tot}} + D_f \textbf{u}_{\text{tot}}
\end{array}
\end{equation}
where
\begin{equation*}
\textbf{x}_{\text{tot}} = \begin{bmatrix}
\dot{V}_x\\V_x\\V_y\\\dot{\psi}\\e_1\\e_2
\end{bmatrix},
\qquad
\textbf{u}_{\text{tot}} = 
\begin{bmatrix}
a\\\delta\\\dot{\psi}_{\text{des}}
\end{bmatrix}
\end{equation*}
and

\[ 
A_f=\begin{bmatrix}
A_m&\textbf{0}_{2\times4}\\
\textbf{0}_{4\times2}&A_a
\end{bmatrix},
\quad
B_f=\begin{bmatrix}
B_m&\textbf{0}_{2\times2}\\
\textbf{0}_{4\times1}&B_a
\end{bmatrix},
\]
\[
C_f=\begin{bmatrix}
C_m&\textbf{0}_{1\times4}\\
\textbf{0}_{2\times2}&C_a
\end{bmatrix}, 
\quad
D_f=\textbf{0}_{3\times3}. 
\]

@@

	\bibitem{long_tf}Y. Wang, Y. Bin and K. Li, "Longitudinal acceleration tracking control of low speed heavy-duty vehicles," in Tsinghua Science and Technology, vol. 13, no. 5, pp. 636-643, Oct. 2008.

@@

	\bibitem{rathai}Murali Madhavan Rathai, Karthik \& Amirthalingam, Jegan \& Jayaraman, Balaji. (2017). Robust tube-MPC based lane keeping system for autonomous driving vehicles. 1-6.
